---
title: "ST558 Fall 2023, Project 2"
author: "Terry McTest"
date: "2023-10-11"
output: github_document
---
<br>

This vignette focuses on a custom function that was developed for use in conjunction with a particular API (Application Programming Interface) in order to query, parse, and return well-structured data.  After using said function to obtain data, an exploratory data analysis will be performed.
<br>


# The Official Carbon Intensity API for Great Britain

The Carbon Intensity API provides information regarding regional carbon intensity of the electricity system in Great Britain.  Since carbon intensity varies by hour, day, and season (due changes in demand as well as variation in the means of generating electricity), these data are stored at the micro-level, e.g. individual carbon intensity readings for each half-hour period over a number of years. Per the [National Grid’s Carbon Intensity API website] (https://carbon-intensity.github.io/api-definitions/?shell#carbon-intensity-api-v2-0-0):

>National Grid’s Carbon Intensity API provides an indicative trend of regional carbon intensity of the electricity system in Great Britain (GB) up to 2 days ahead of real-time. It provides programmatic and timely access to both forecast and estimated carbon intensity data. The Carbon Intensity forecast includes CO2 emissions related to electricity generation only. The includes emissions from all large metered power stations, interconnector imports, transmission and distribution losses, and accounts for national electricity demand, embedded wind and solar generation.

The API consists of three primary endpoints:

* *National-level carbon intensity*.  This endpoint includes a continuous measure of forecasted CO2 emissions per kilowatt hour of electricity produced; a continuous measure of actual CO2 emissions produced; and a categorical measure of actual CO2 emissions produced. 
* *National-level electricity-generation mix*.  This endpoint includes the percent distribution of how electricity is generated across various generation sources, including biomass, coal, imports, gas, nuclear, hydro, solar, wind, and "other".
* *Regional-level carbon intensity & electricity-generation mix*.  This endpoint includes (most) measures made available in the above endpoints, though measures in the regional-level endpoint are stored at a regional level (i.e. for each of 18 regions of Great Britain) as opposed to being stored at a national level.
<br>



# Function for Returning Data From the API

Before proceeding, we should make note of the specific R packages that are employed by either the custom function presented here, and/or by the exploratory data analysis to follow:
```{r, message=FALSE}
library(httr)
library(dplyr)
library(tidyr)
library(jsonlite)
library(tidyverse)
library(ggradar)
library(scales)
```
<br>

A function intended to return data from the Carbon Intensity (CI) API must be capable of navigating certain features of this API:

* The CI API needs to know the window of time for which data is desired, even if that "window" is essentially a specific point in time (e.g. "right now") as opposed to an actual extended window.
* The CI API is comprised of data objects nested within data objects (e.g. data frames within data frames), and these structures must be un-nested and re-structured to facilitate use of the data.
* In general terms, the national data is structured as one record per each half-hour period contained within the specified window of time, while the regional data is structured with 18 records (one for each region) per each half-hour period contained within the specified window of time. The function must account for whether one or both types of data are desired (i.e. national-level data and/or regional-level data).

The function presented here addresses the above features of the CI API by first requesting specification of the following arguments:

* `from_dt` and `to_dt`:  These are required arguments, each representing a timestamp which bookends the window of time for which data is desired.
* `want_int`, `want_gen`, `want_reg_wide`, and `want_reg_long`: These are optional arguments through which the end user specifies (by setting the applicable argument equal to 1) whether they are requesting national-level intensity data; national-level generation-mix data; regional-level data (in wide format); and/or regional-level data (in long format).

The output of the CI API function call (named simply `carbon` as shown here) is a list comprised of either: one wide dataset + one long dataset; a wide dataset only; or a long dataset only.  More specifically, by selectively setting the aforementioned optional arguments to 1, the user will receive a wide dataset comprised of either:

1. national-level intensity data + national-level generation-mix data + regional-level data;
2. national-level intensity data + national-level generation-mix data;
3. national-level intensity data + regional-level data;
4. national-level generation-mix data + regional-level data;
5. national-level intensity data only;
6. national-level generation-mix data only; or
7. regional-level data only

Each version of the wide dataset listed above can be output with or without an accompanying long dataset of regional-level data, for a total of (7 x 2) + 1 = 15 data request options/combinations (the "+1" option would be the long dataset of regional-level data with no wide dataset).  Notably, each of these 15 options/combinations could be used in conjunction with an infinite number of time windows, providing the user with an even greater number of options for each data request.  

As seen below, the `carbon` function also performs some niceties such as imposing order on categorical variables, and converting the bookend timestamps (which are actually character textstrings in "YYYY-MM-DDTHH:MMZ" format as required to access the CI API) to various date values to facilitate the forthcoming exploratory data analysis.

The code chunk to follow shows R syntax for the entire `carbon` function, annotated with comments to describe the purpose of each sub-chunk of code:
```{r}
carbon <- function(from_dt, to_dt, want_int=0, want_gen=0, want_reg_wide=0, want_reg_long=0)

{
  
#national-level carbon intensity data/endpoint

if(want_int==1)
{
base = "https://api.carbonintensity.org.uk/intensity/"
slash = "/"
int_url = paste0(base, from_dt, slash, to_dt)

int_api <- GET(int_url)
parsed_int <- fromJSON(rawToChar(int_api$content))
#str(parsed_int$data)
#as_tibble(parsed_int$data)

int <-
  parsed_int$data %>%
  as_tibble() %>%
  unnest(intensity)
}

  

#national-level electricity-generation mix data/endpoint

if(want_gen==1)
{
base = "https://api.carbonintensity.org.uk/generation/"
slash = "/"
gen_url = paste0(base, from_dt, slash, to_dt)
  
gen_api <- GET(gen_url)
parsed_gen <- fromJSON(rawToChar(gen_api$content))
#str(parsed_gen$data)
#as_tibble(parsed_gen$data)

gen <-
  parsed_gen$data %>%
  as_tibble() %>%
  unnest(generationmix) %>%
  pivot_wider(names_from = fuel, values_from = perc)
}


  
#regional-level data/endpoint
  
if(want_reg_wide==1 | want_reg_long==1)
{
base = "https://api.carbonintensity.org.uk/regional/intensity/"
slash = "/"
reg_url = paste0(base, from_dt, slash, to_dt)
  
regional_api <- GET(reg_url)
parsed_regional <- fromJSON(rawToChar(regional_api$content))
#str(parsed_regional$data)
#as_tibble(parsed_regional$data)
  
reg_long <-
  parsed_regional$data %>%
  as_tibble() %>%
  unnest(regions) %>%
  unnest(intensity) %>%
  unnest(generationmix) %>%
  pivot_wider(names_from = fuel, values_from = perc)
}
  
if(want_reg_wide==1)
{
reg_wide <-
 reg_long %>%
  select(from, to, regionid, forecast, index, biomass, coal, imports, gas, nuclear, other, hydro, solar, wind) %>%
  pivot_wider(names_from = regionid, 
    values_from = c(forecast, index, biomass, coal, imports, gas, nuclear, other, hydro, solar, wind))
}
  
  

#compile the *wide* dataset for outputting
  
if(want_int==1 & want_gen==1 & want_reg_wide==1)
  {carbon_wide <- full_join(int, gen, by=c("to"="to", "from"="from"))
   carbon_wide <- full_join(carbon_wide, reg_wide, by=c("to"="to", "from"="from"))}
  
else if(want_int==1 & want_gen==1)
  {carbon_wide <- full_join(int, gen, by=c("to"="to", "from"="from"))}

else if(want_int==1 & want_reg_wide==1)
  {carbon_wide <- full_join(int, reg_wide, by=c("to"="to", "from"="from"))}
  
else if(want_gen==1 & want_reg_wide==1)
  {carbon_wide <- full_join(gen, reg_wide, by=c("to"="to", "from"="from"))}

else if(want_int==1)
  {carbon_wide <- int}
  
else if(want_gen==1)
  {carbon_wide <- gen}

else if(want_reg_wide==1)
  {carbon_wide <- reg_wide}


  

#to facilitate downstream analyses, derive a set of more-usable date/season vars

if(want_int==1 | want_gen==1 | want_reg_wide==1)
{
carbon_wide$yyyy = substring(carbon_wide$from, first=1, last=4)
carbon_wide$mm = substring(carbon_wide$from, first=6, last=7)
carbon_wide$yyyymm = paste0(carbon_wide$yyyy, carbon_wide$mm)
carbon_wide$date = as.Date(substring(carbon_wide$from, first=1, last=10), format="%Y-%m-%d")

carbon_wide$season =
  if_else(carbon_wide$mm %in% c("04", "05", "06"), "spring",
  if_else(carbon_wide$mm %in% c("07", "08", "09"), "summer",
  if_else(carbon_wide$mm %in% c("10", "11", "12"), "fall",
  if_else(carbon_wide$mm %in% c("01", "02", "03"), "winter", NA))))
}

if(want_reg_long==1)
{
reg_long$yyyy = substring(reg_long$from, first=1, last=4)
reg_long$mm = substring(reg_long$from, first=6, last=7)
reg_long$yyyymm = paste0(reg_long$yyyy, reg_long$mm)
reg_long$date = as.Date(substring(reg_long$from, first=1, last=10), format="%Y-%m-%d")

reg_long$season =
  if_else(reg_long$mm %in% c("04", "05", "06"), "spring",
  if_else(reg_long$mm %in% c("07", "08", "09"), "summer",
  if_else(reg_long$mm %in% c("10", "11", "12"), "fall",
  if_else(reg_long$mm %in% c("01", "02", "03"), "winter", NA))))
}
  
  
  
#impose ordering on categorical variables
  
if(want_int==1 | want_gen==1 | want_reg_wide==1)
{
carbon_wide$index <- ordered(carbon_wide$index, levels = c("very high", "high", "moderate", "low", "very low"))
carbon_wide$season <- ordered(carbon_wide$season, levels = c("spring", "summer", "fall", "winter"))
}
  
if(want_reg_long==1)
{
reg_long$index <- ordered(reg_long$index, levels = c("very high", "high", "moderate", "low", "very low"))
reg_long$season <- ordered(reg_long$season, levels = c("spring", "summer", "fall", "winter"))
}  
  

  
    
#final output, which is either:
#(1) a list comprised of one long dataset + one wide dataset, (2) a single wide dataset, or (3) a single long dataset
# (the long dataset will always be regional data)
  
if(want_reg_long==1 & (want_int==1 | want_gen==1 | want_reg_wide==1))
  {carbon_dat <- list(reg_long=reg_long, carbon_wide=carbon_wide)}
  
else if(want_int==1 | want_gen==1 | want_reg_wide==1)
  {carbon_dat <- carbon_wide}

else if(want_reg_long==1)
  {carbon_dat <- reg_long}
  
}
```
<br>



# Using the Custom Function to Return Data
```{r}
jan2019 <- carbon(from_dt="2019-01-08T12:00Z", to_dt="2019-01-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
feb2019 <- carbon(from_dt="2019-02-08T12:00Z", to_dt="2019-02-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
mar2019 <- carbon(from_dt="2019-03-08T12:00Z", to_dt="2019-03-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
apr2019 <- carbon(from_dt="2019-04-08T12:00Z", to_dt="2019-04-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
may2019 <- carbon(from_dt="2019-05-08T12:00Z", to_dt="2019-05-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
jun2019 <- carbon(from_dt="2019-06-08T12:00Z", to_dt="2019-06-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
jul2019 <- carbon(from_dt="2019-07-08T12:00Z", to_dt="2019-07-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
aug2019 <- carbon(from_dt="2019-08-08T12:00Z", to_dt="2019-08-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
sep2019 <- carbon(from_dt="2019-09-08T12:00Z", to_dt="2019-09-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
oct2019 <- carbon(from_dt="2019-10-08T12:00Z", to_dt="2019-10-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
nov2019 <- carbon(from_dt="2019-11-08T12:00Z", to_dt="2019-11-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
dec2019 <- carbon(from_dt="2019-12-08T12:00Z", to_dt="2019-12-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)

jan2020 <- carbon(from_dt="2020-01-08T12:00Z", to_dt="2020-01-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
feb2020 <- carbon(from_dt="2020-02-08T12:00Z", to_dt="2020-02-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
mar2020 <- carbon(from_dt="2020-03-08T12:00Z", to_dt="2020-03-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
apr2020 <- carbon(from_dt="2020-04-08T12:00Z", to_dt="2020-04-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
may2020 <- carbon(from_dt="2020-05-08T12:00Z", to_dt="2020-05-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
jun2020 <- carbon(from_dt="2020-06-08T12:00Z", to_dt="2020-06-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
jul2020 <- carbon(from_dt="2020-07-08T12:00Z", to_dt="2020-07-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
aug2020 <- carbon(from_dt="2020-08-08T12:00Z", to_dt="2020-08-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
sep2020 <- carbon(from_dt="2020-09-08T12:00Z", to_dt="2020-09-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
oct2020 <- carbon(from_dt="2020-10-08T12:00Z", to_dt="2020-10-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
nov2020 <- carbon(from_dt="2020-11-08T12:00Z", to_dt="2020-11-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
dec2020 <- carbon(from_dt="2020-12-08T12:00Z", to_dt="2020-12-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)

jan2021 <- carbon(from_dt="2021-01-08T12:00Z", to_dt="2021-01-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
feb2021 <- carbon(from_dt="2021-02-08T12:00Z", to_dt="2021-02-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
mar2021 <- carbon(from_dt="2021-03-08T12:00Z", to_dt="2021-03-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
apr2021 <- carbon(from_dt="2021-04-08T12:00Z", to_dt="2021-04-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
may2021 <- carbon(from_dt="2021-05-08T12:00Z", to_dt="2021-05-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
jun2021 <- carbon(from_dt="2021-06-08T12:00Z", to_dt="2021-06-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
jul2021 <- carbon(from_dt="2021-07-08T12:00Z", to_dt="2021-07-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
aug2021 <- carbon(from_dt="2021-08-08T12:00Z", to_dt="2021-08-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
sep2021 <- carbon(from_dt="2021-09-08T12:00Z", to_dt="2021-09-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
oct2021 <- carbon(from_dt="2021-10-08T12:00Z", to_dt="2021-10-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
nov2021 <- carbon(from_dt="2021-11-08T12:00Z", to_dt="2021-11-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
dec2021 <- carbon(from_dt="2021-12-08T12:00Z", to_dt="2021-12-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)

jan2022 <- carbon(from_dt="2022-01-08T12:00Z", to_dt="2022-01-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
feb2022 <- carbon(from_dt="2022-02-08T12:00Z", to_dt="2022-02-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
mar2022 <- carbon(from_dt="2022-03-08T12:00Z", to_dt="2022-03-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
apr2022 <- carbon(from_dt="2022-04-08T12:00Z", to_dt="2022-04-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
may2022 <- carbon(from_dt="2022-05-08T12:00Z", to_dt="2022-05-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
jun2022 <- carbon(from_dt="2022-06-08T12:00Z", to_dt="2022-06-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
jul2022 <- carbon(from_dt="2022-07-08T12:00Z", to_dt="2022-07-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
aug2022 <- carbon(from_dt="2022-08-08T12:00Z", to_dt="2022-08-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
sep2022 <- carbon(from_dt="2022-09-08T12:00Z", to_dt="2022-09-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
oct2022 <- carbon(from_dt="2022-10-08T12:00Z", to_dt="2022-10-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
nov2022 <- carbon(from_dt="2022-11-08T12:00Z", to_dt="2022-11-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)
dec2022 <- carbon(from_dt="2022-12-08T12:00Z", to_dt="2022-12-21T12:00Z", want_int=1, want_gen=1, want_reg_wide=1)


wide <-
  bind_rows(
  jan2019, feb2019, mar2019, apr2019, may2019, jun2019, jul2019, aug2019, sep2019, oct2019, nov2019, dec2019,
  jan2020, feb2020, mar2020, apr2020, may2020, jun2020, jul2020, aug2020, sep2020, oct2020, nov2020, dec2020,
  jan2021, feb2021, mar2021, apr2021, may2021, jun2021, jul2021, aug2021, sep2021, oct2021, nov2021, dec2021,
  jan2022, feb2022, mar2022, apr2022, may2022, jun2022, jul2022, aug2022, sep2022, oct2022, nov2022, dec2022)
```
<br>


```{r}
jan2019 <- carbon(from_dt="2019-01-08T12:00Z", to_dt="2019-01-21T12:00Z", want_reg_long=1)
feb2019 <- carbon(from_dt="2019-02-08T12:00Z", to_dt="2019-02-21T12:00Z", want_reg_long=1)
mar2019 <- carbon(from_dt="2019-03-08T12:00Z", to_dt="2019-03-21T12:00Z", want_reg_long=1)
apr2019 <- carbon(from_dt="2019-04-08T12:00Z", to_dt="2019-04-21T12:00Z", want_reg_long=1)
may2019 <- carbon(from_dt="2019-05-08T12:00Z", to_dt="2019-05-21T12:00Z", want_reg_long=1)
jun2019 <- carbon(from_dt="2019-06-08T12:00Z", to_dt="2019-06-21T12:00Z", want_reg_long=1)
jul2019 <- carbon(from_dt="2019-07-08T12:00Z", to_dt="2019-07-21T12:00Z", want_reg_long=1)
aug2019 <- carbon(from_dt="2019-08-08T12:00Z", to_dt="2019-08-21T12:00Z", want_reg_long=1)
sep2019 <- carbon(from_dt="2019-09-08T12:00Z", to_dt="2019-09-21T12:00Z", want_reg_long=1)
oct2019 <- carbon(from_dt="2019-10-08T12:00Z", to_dt="2019-10-21T12:00Z", want_reg_long=1)
nov2019 <- carbon(from_dt="2019-11-08T12:00Z", to_dt="2019-11-21T12:00Z", want_reg_long=1)
dec2019 <- carbon(from_dt="2019-12-08T12:00Z", to_dt="2019-12-21T12:00Z", want_reg_long=1)

jan2020 <- carbon(from_dt="2020-01-08T12:00Z", to_dt="2020-01-21T12:00Z", want_reg_long=1)
feb2020 <- carbon(from_dt="2020-02-08T12:00Z", to_dt="2020-02-21T12:00Z", want_reg_long=1)
mar2020 <- carbon(from_dt="2020-03-08T12:00Z", to_dt="2020-03-21T12:00Z", want_reg_long=1)
apr2020 <- carbon(from_dt="2020-04-08T12:00Z", to_dt="2020-04-21T12:00Z", want_reg_long=1)
may2020 <- carbon(from_dt="2020-05-08T12:00Z", to_dt="2020-05-21T12:00Z", want_reg_long=1)
jun2020 <- carbon(from_dt="2020-06-08T12:00Z", to_dt="2020-06-21T12:00Z", want_reg_long=1)
jul2020 <- carbon(from_dt="2020-07-08T12:00Z", to_dt="2020-07-21T12:00Z", want_reg_long=1)
aug2020 <- carbon(from_dt="2020-08-08T12:00Z", to_dt="2020-08-21T12:00Z", want_reg_long=1)
sep2020 <- carbon(from_dt="2020-09-08T12:00Z", to_dt="2020-09-21T12:00Z", want_reg_long=1)
oct2020 <- carbon(from_dt="2020-10-08T12:00Z", to_dt="2020-10-21T12:00Z", want_reg_long=1)
nov2020 <- carbon(from_dt="2020-11-08T12:00Z", to_dt="2020-11-21T12:00Z", want_reg_long=1)
dec2020 <- carbon(from_dt="2020-12-08T12:00Z", to_dt="2020-12-21T12:00Z", want_reg_long=1)

jan2021 <- carbon(from_dt="2021-01-08T12:00Z", to_dt="2021-01-21T12:00Z", want_reg_long=1)
feb2021 <- carbon(from_dt="2021-02-08T12:00Z", to_dt="2021-02-21T12:00Z", want_reg_long=1)
mar2021 <- carbon(from_dt="2021-03-08T12:00Z", to_dt="2021-03-21T12:00Z", want_reg_long=1)
apr2021 <- carbon(from_dt="2021-04-08T12:00Z", to_dt="2021-04-21T12:00Z", want_reg_long=1)
may2021 <- carbon(from_dt="2021-05-08T12:00Z", to_dt="2021-05-21T12:00Z", want_reg_long=1)
jun2021 <- carbon(from_dt="2021-06-08T12:00Z", to_dt="2021-06-21T12:00Z", want_reg_long=1)
jul2021 <- carbon(from_dt="2021-07-08T12:00Z", to_dt="2021-07-21T12:00Z", want_reg_long=1)
aug2021 <- carbon(from_dt="2021-08-08T12:00Z", to_dt="2021-08-21T12:00Z", want_reg_long=1)
sep2021 <- carbon(from_dt="2021-09-08T12:00Z", to_dt="2021-09-21T12:00Z", want_reg_long=1)
oct2021 <- carbon(from_dt="2021-10-08T12:00Z", to_dt="2021-10-21T12:00Z", want_reg_long=1)
nov2021 <- carbon(from_dt="2021-11-08T12:00Z", to_dt="2021-11-21T12:00Z", want_reg_long=1)
dec2021 <- carbon(from_dt="2021-12-08T12:00Z", to_dt="2021-12-21T12:00Z", want_reg_long=1)

jan2022 <- carbon(from_dt="2022-01-08T12:00Z", to_dt="2022-01-21T12:00Z", want_reg_long=1)
feb2022 <- carbon(from_dt="2022-02-08T12:00Z", to_dt="2022-02-21T12:00Z", want_reg_long=1)
mar2022 <- carbon(from_dt="2022-03-08T12:00Z", to_dt="2022-03-21T12:00Z", want_reg_long=1)
apr2022 <- carbon(from_dt="2022-04-08T12:00Z", to_dt="2022-04-21T12:00Z", want_reg_long=1)
may2022 <- carbon(from_dt="2022-05-08T12:00Z", to_dt="2022-05-21T12:00Z", want_reg_long=1)
jun2022 <- carbon(from_dt="2022-06-08T12:00Z", to_dt="2022-06-21T12:00Z", want_reg_long=1)
jul2022 <- carbon(from_dt="2022-07-08T12:00Z", to_dt="2022-07-21T12:00Z", want_reg_long=1)
aug2022 <- carbon(from_dt="2022-08-08T12:00Z", to_dt="2022-08-21T12:00Z", want_reg_long=1)
sep2022 <- carbon(from_dt="2022-09-08T12:00Z", to_dt="2022-09-21T12:00Z", want_reg_long=1)
oct2022 <- carbon(from_dt="2022-10-08T12:00Z", to_dt="2022-10-21T12:00Z", want_reg_long=1)
nov2022 <- carbon(from_dt="2022-11-08T12:00Z", to_dt="2022-11-21T12:00Z", want_reg_long=1)
dec2022 <- carbon(from_dt="2022-12-08T12:00Z", to_dt="2022-12-21T12:00Z", want_reg_long=1)


long <-
  bind_rows(
  jan2019, feb2019, mar2019, apr2019, may2019, jun2019, jul2019, aug2019, sep2019, oct2019, nov2019, dec2019,
  jan2020, feb2020, mar2020, apr2020, may2020, jun2020, jul2020, aug2020, sep2020, oct2020, nov2020, dec2020,
  jan2021, feb2021, mar2021, apr2021, may2021, jun2021, jul2021, aug2021, sep2021, oct2021, nov2021, dec2021,
  jan2022, feb2022, mar2022, apr2022, may2022, jun2022, jul2022, aug2022, sep2022, oct2022, nov2022, dec2022)
```
<br>



# Exploratory Data Analysis

```{r}
#national, over time
ggplot(wide, aes(x = date, y = actual)) + geom_line() + geom_smooth() +
labs(y = "CO2 emissions (in gCO2)",
     title = "CO2 Emissions in Great Britain: 2019-2022")
  
wide %>%
  group_by(yyyy) %>%
  summarize(mean_actual = mean(actual, na.rm=TRUE))
```


```{r}
#scatterplot, forecast vs. actual
ggplot(wide, aes(x = forecast, y = actual)) + geom_point() +
  labs(y = "actual CO2 emissions (in gCO2)", 
       x = "forecast CO2 emissions (in gCO2)", 
       title = "Forecast vs. Actual CO2 Emissions (outliers included): 2019-2022")

no_outliers <- wide %>% filter(forecast < 1000)
ggplot(no_outliers, aes(x = forecast, y = actual)) + geom_point() +
labs(y = "actual CO2 emissions (in gCO2)", 
     x = "forecast CO2 emissions (in gCO2)", 
     title = "Forecast vs. Actual CO2 Emissions (outliers excluded): 2019-2022")
```


```{r}
#by season
wide$index <- ordered(wide$index, levels = c("very high", "high", "moderate", "low", "very low"))
wide$season <- ordered(wide$season, levels = c("spring", "summer", "fall", "winter"))
#can do the above ordering in the function

ggplot(wide, aes(x=season)) + geom_bar(aes(fill = as.factor(index)), position = "dodge") + 
  labs(title = "Counts of Daily CO2 Emissions Indices, by Season: 2019-2022") +
  scale_fill_discrete(name = "Emissions Index")

table(wide$season, wide$index)

wide %>%
  group_by(season) %>%
  summarize(mean = mean(actual, na.rm=TRUE))
```


```{r}
#over time, by region
ggplot(long, aes(x = date, y = forecast, color=shortname)) + geom_smooth() +
  labs(y = "forecast CO2 emissions (in gCO2)", 
       title = "Forecast CO2 Emissions (as a proxy for actual emissions), by Region: 2019-2022") +
  scale_color_discrete(name = "Region")  

#limit to principle regions
e_s_w <-
  long %>%
  filter(shortname %in% c("England", "Scotland", "Wales", "GB"))

ggplot(e_s_w, aes(x = date, y = forecast, color=shortname)) + geom_smooth() +
  labs(y = "forecast CO2 emissions (in gCO2)", 
       title = "Forecast CO2 Emissions (as a proxy for actual emissions), by Principal Region: 2019-2022") +
  scale_color_discrete(name = "Region")
```


```{r}
#association between source of electricity and CO2 emissions
ggplot(wide, aes(x = biomass, y = actual, position = "jitter")) + 
  geom_point() + geom_smooth() +
  labs(y = "actual CO2 emissions (in gCO2)", 
     x = "proportion of electricity generated by biomass", 
     title = "Biomass-Generated Electricity vs. CO2 Emissions: 2019-2022")

ggplot(wide, aes(x = coal, y = actual, position = "jitter")) + 
  geom_point() + geom_smooth() +
  labs(y = "actual CO2 emissions (in gCO2)", 
       x = "proportion of electricity generated by coal", 
       title = "Coal-Generated Electricity vs. CO2 Emissions: 2019-2022")

ggplot(wide, aes(x = imports, y = actual, position = "jitter")) + 
  geom_point() + geom_smooth() +
  labs(y = "actual CO2 emissions (in gCO2)", 
       x = "proportion of electricity that is imported", 
       title = "Imported Electricity vs. CO2 Emissions: 2019-2022")

ggplot(wide, aes(x = gas, y = actual, position = "jitter")) + 
  geom_point() + geom_smooth() +
  labs(y = "actual CO2 emissions (in gCO2)", 
       x = "proportion of electricity generated by gas", 
       title = "Gas-Generated Electricity vs. CO2 Emissions: 2019-2022")

ggplot(wide, aes(x = nuclear, y = actual, position = "jitter")) + 
  geom_point() + geom_smooth() +
  labs(y = "actual CO2 emissions (in gCO2)", 
       x = "proportion of electricity that is nuclear-generated", 
       title = "Nuclear-Generated Electricity vs. CO2 Emissions: 2019-2022")

ggplot(wide, aes(x = hydro, y = actual, position = "jitter")) + 
  geom_point() + geom_smooth() +
  labs(y = "actual CO2 emissions (in gCO2)", 
       x = "proportion of electricity that is hyrdro-generated", 
       title = "Hydro-Generated Electricity vs. CO2 Emissions: 2019-2022")

ggplot(wide, aes(x = solar, y = actual, position = "jitter")) + 
  geom_point() + geom_smooth() +
  labs(y = "actual CO2 emissions (in gCO2)", 
       x = "proportion of electricity that is solar-generated", 
       title = "Solar-Generated Electricity vs. CO2 Emissions: 2019-2022")

ggplot(wide, aes(x = wind, y = actual, position = "jitter")) + 
  geom_point() + geom_smooth() +
  labs(y = "actual CO2 emissions (in gCO2)", 
       x = "proportion of electricity that is wind-generated", 
       title = "Wind-Generated Electricity vs. CO2 Emissions: 2019-2022")
```


```{r}
# generation mix
by_year <-
  wide %>%
  group_by(yyyy) %>%
  summarize(biomass = mean(biomass, na.rm=TRUE),
            coal = mean(coal, na.rm=TRUE),
            imports = mean(imports, na.rm=TRUE),
            gas = mean(gas, na.rm=TRUE),
            nuclear = mean(nuclear, na.rm=TRUE),
            other = mean(other, na.rm=TRUE),
            hydro = mean(hydro, na.rm=TRUE),
            solar = mean(solar, na.rm=TRUE),
            wind = mean(wind, na.rm=TRUE))

ggplot(by_year, aes(x=yyyy, group=1)) +
  geom_line(aes(y=biomass, color="biomass")) +
  geom_line(aes(y=coal, color="coal")) +
  geom_line(aes(y=imports, color="imports")) +
  geom_line(aes(y=gas, color="gas")) +
  geom_line(aes(y=nuclear, color="nuclear")) +
  geom_line(aes(y=hydro, color="hydro")) +
  geom_line(aes(y=solar, color="solar")) +
  geom_line(aes(y=wind, color="wind")) +
  geom_line(aes(y=other, color="other")) +
  scale_color_discrete(name = "generation source") +
  labs(y = "Percentage",
    title = "Percentage of Total Electricity Generated, by Source (yearly averages)")

ggplot(wide, aes(x=date)) +
  geom_smooth(aes(y=biomass, color="biomass")) +
  geom_smooth(aes(y=coal, color="coal")) +
  geom_smooth(aes(y=imports, color="imports")) +
  geom_smooth(aes(y=gas, color="gas")) +
  geom_smooth(aes(y=nuclear, color="nuclear")) +
  geom_smooth(aes(y=hydro, color="hydro")) +
  geom_smooth(aes(y=solar, color="solar")) +
  geom_smooth(aes(y=wind, color="wind")) +
  geom_smooth(aes(y=other, color="other"))
```


```{r}
#radar plot of generation mix by region
library(ggradar)
library(scales)
reg_radar <-
  long %>%
  drop_na() %>%
  filter(shortname %in% c("England", "Scotland", "Wales", "GB")) %>%
  group_by(shortname) %>%
  summarize(biomass = mean(biomass, na.rm=TRUE),
            coal = mean(coal, na.rm=TRUE),
            imports = mean(imports, na.rm=TRUE),
            gas = mean(gas, na.rm=TRUE),
            nuclear = mean(nuclear, na.rm=TRUE),
            #other = mean(other, na.rm=TRUE),
            hydro = mean(hydro, na.rm=TRUE),
            solar = mean(solar, na.rm=TRUE),
            wind = mean(wind, na.rm=TRUE)) %>%
  ungroup() %>%
  mutate_at(vars(-shortname), rescale)

ggradar(reg_radar) + labs(title = "Radar Plot of Electricity-Generation Mix, by Region") 

  #above code is based on https://r-graph-gallery.com/web-radar-chart-with-R.html
```


```{r}
#actual (i.e. un-scaled) proportions underlying the radar plot
long %>%
  filter(shortname %in% c("England", "Scotland", "Wales", "GB")) %>%
  group_by(shortname) %>%
  summarize(biomass = mean(biomass, na.rm=TRUE),
            coal = mean(coal, na.rm=TRUE),
            imports = mean(imports, na.rm=TRUE),
            gas = mean(gas, na.rm=TRUE),
            nuclear = mean(nuclear, na.rm=TRUE),
            hydro = mean(hydro, na.rm=TRUE),
            solar = mean(solar, na.rm=TRUE),
            wind = mean(wind, na.rm=TRUE),
            other = mean(other, na.rm=TRUE))
```